Computing Entropies With Nested Sampling

### 2017-07-10 paper/ms.tex
Bit more work

Consider the basic statistics problem of inferring a quantity $\mu$ from
100 observations $\d = \{d_1, ..., d_{100}\}$ whose
probability distribution (conditional on $\mu$) is
\begin{align}
p(d_i | \mu) &\sim \textnormal{Normal}(\mu, 1).
\end{align}
If the prior for $\mu$ is Normal$(0, 10^2)$, then the posterior is
\begin{align}
\mu | \d &\sim \textnormal{Normal}\left(
                                       \frac{1}{100}\sum_{i=1}^{100} d_i,
                                       \left[\frac{10}{\sqrt{10001}}\right]^2
                                       \right).
\end{align}
The true value of the differential entropy $H(\d)$ is available analytically
in this case, and its value is $146.499$ nats. However, in most situations, the
entropy of a marginal distribution is not available in closed form.

I ran the algorithm and computed the depth using a tolerance of
$L=10^{-3}\sqrt{100}$, so that the RMS difference between
points in the two datasets was about $10^{-3}$.
From the algorithm, the average depth was
\begin{align}
\depth_{\d} &= 698.25 \pm 0.34.
\end{align}
The log-volume of a 100-dimensional ball of radius
$L=10^{-3}\sqrt{100}$ is
$-551.76$. Therefore, the differential entropy is estimated to be
\begin{align}
H(\d) &\approx 146.49 \pm 0.34.
\end{align}
which is very close to the true value (suspiciously close, but this
was just a fluke).

>>>>

Consider the basic statistics problem of inferring a quantity $\mu$ from
100 observations $\boldsymbol{x} = \{x_1, ..., x_{100}\}$ whose
probability distribution (conditional on $\mu$) is
\begin{align}
p(x_i | \mu) &\sim \textnormal{Normal}(\mu, 1).
\end{align}
If the prior for $\mu$ is Normal$(0, 10^2)$, then the posterior is
\begin{align}
\mu | \boldsymbol{x} &\sim \textnormal{Normal}\left(
                                       \frac{1}{100}\sum_{i=1}^{100} x_i,
                                       \left[\frac{10}{\sqrt{10001}}\right]^2
                                       \right).
\end{align}
The true value of the differential entropy $H(\boldsymbol{x})$ is
available analytically
in this case, and its value is $146.499$ nats. However, in most situations, the
entropy of a marginal distribution is not available in closed form.

To use the algorithm, a distance function $d(,)$ must be defined, to quantify
how far the NS particles are from the reference particle. To calculate
the entropy of the data, I used the following distance function:
\begin{align}
d\left(\x; \xref\right) &= \sum_{i=1}^{100}\left(x_i - x_{{\rm ref}, i}\right)^2
\end{align}

I ran the algorithm and computed the depth using a tolerance of
$L=10^{-3}\sqrt{100}$, so that the RMS difference between
points in the two datasets was about $10^{-3}$.
From the algorithm, the average depth was
\begin{align}
\depth_{\boldsymbol{x}} &= 698.25 \pm 0.34.
\end{align}
The log-volume of a 100-dimensional ball of radius
$L=10^{-3}\sqrt{100}$ is
$-551.76$. Therefore, the differential entropy is estimated to be
\begin{align}
H(\boldsymbol{x}) &\approx 146.49 \pm 0.34.
\end{align}
which is very close to the true value (suspiciously close, but this
was just a fluke).
